Extended Formal Analysis of the EDHOC Protocol in Tamarin

### 2022-08-09 paper.tex
[add] clarifications and spelling corrections

%-------------------------------------------------------------------------- sub
\subsection{Contributions}
\label{sec:contributions}
In this paper, we extend the work presented in~\cite{Norr21}.
%
We formally analyze the \mEdhoc{} specification as of July
2020~\cite{our-analysis-selander-lake-edhoc-00}, but our formal analysis applies as far as the version of the \mSpec{} from February 2021.
%
We extend the adversary model and \mTamarin{} system models to capture weak
post-compromise security (PCS) and model Trusted Execution Environments
(TEE).
%
We also alter the formal modelling used for encryption under \mXor{}.
%
We formally verify the following properties:
\begin{itemize}
\item Injective agreement
\item Implicit agreement
\item Perfect Forward Secrecy (PFS) for the session-key material
\item Weak post-compromise security for the session-key material
\item Secrecy and integrity of \mADthree
\end{itemize}
%
We follow the definition of weak PCS by~\cite{DBLP:conf/csfw/Cohn-GordonCG16},
which subsumes PFS.
%
Since \mEdhoc{} is still a protocol under development, the latest version of \mEdhoc{} is the one from July 2022~\cite{draft-ietf-lake-edhoc-15}, which contains details which are not covered by our formal model.
%
We do, however, refer to various aspects
(error handling, denial of service etc) of the July 2022 version of the \mSpec{}, namely 
in Sections~\ref{sec:errorHandling} and \ref{sec:discussion}.
%
We also discuss various issues arising due to the use of trusted
execution environments (TEEs), denial of service (DoS) attacks, error handling,
the negotiation of parameters (which the formal model abstracts
away) for establishing the protocol, and other potential attacks and concerns.
%
We have communicated these issues to the developers of the protocol.
%

@@

%-------------------------------------------------------------------------- sub
\subsection{Session Key Material}
\label{sec:sessionKeyMaterial}
As we have seen, \mEdhoc{} establishes some
session-key material, which can be fed into the \mEdhoc{}-Exporter to derive
session keys.
%
This key material is directly influenced by \mGxy{} and a party's secret static
long-term key, if they use \mStat{}.
%
As mentioned earlier, mutual explicit injective agreement cannot be obtained for
$P_{I}$.
%
This might prove to be a problem, and we proposed three potential alternatives
for addressing this to the \mEdhoc{} authors.
%
\begin{itemize}
    \item Include \mIdcredi{} or a hash thereof in the first and
    second messages:
%
    This increases message size and also leaks the initiator's identity.
%
    \item Derive session-key material without using $P_{I}$:
%
    This deviates from the structure of protocols like \mOptls{} from which the
    \mStat{}-based methods draw heavy inspiration.
%
    In those protocols, including $P_{I}$ while computating session-key
    material is crucial to obtaining resistance against the compromise of the
    initiator's ephemeral key.
%
    \item Include a fourth message from the
    responder to the initiator which contains a MAC based on information
    obtained using $P_{I}$:
\end{itemize}
%
While the last option increases the number of messages in the protocol, the
\mEdhoc{} authors decided to go with this option for situations where it is
necessary to provide explicit injective agreement on $P_{I}$.
%

%-------------------------------------------------------------------------- sec
\section{\uppercase{Conclusions and Future Work}}
\label{sec:conclusion}
As part of this paper, we modeled all four authentication methods of the \mEdhoc{} protocol using the \mTamarin{} verification tool.
%
We used a server with 2 Intel(R) Xeon(R) Gold 6242 CPU @ 2.80GHz, which had a total of 32 cores / 64 threads, and of the total 384GiB, the server allotted 256GiB of memory for the \mTamarin{} job.
%
We report both the time taken to complete the task (``real time'') and the CPU time spent on the task (``user time'').
%
Time measurements are rounded to the nearest second.
%
We formulated and verified the following properties in a precise adversary model with controlled access to TEEs and a precise modelling of \mXor.
%
\begin{itemize}
\item Injective agreement for both $I$ and $R$
\item Implicit agreement for both $I$ and $R$
\item Session key secrecy
\item Secrecy and integrity of \mADthree
\end{itemize}
%
We consider two separate sets of parameters, $S_P$ and $S_F$. 
%
The set $S_P$ consists of the following pieces of information: the identity of the responder, roles, context identifierss \mCi{} and \mCr, cipher suites \mSuites, and session-key material excluding $P_I$.
%
For injective agreement, when the initiator uses a \mStat{} method, they are ensured agreement only on this above set of parameters. 
%
The set $S_F$ is the set $S_P$ along with the initiator's identity and $P_I$. 
%
The responder is guaranteed injective agreement on the parameters $S_F$, irrespective of what method they run, while the initiator is guaranteed injective agreement on $S_F$ as long as they run a \mSig{} method.
%
Implicit agreement can be reached for both parties on the set $S_F$.
%
In addition, mutual entity authentication, UKS- and KCI-resistance can be inferred from the verified properties.
%
We present the results in Table~\ref{tab:props}.

%We formally modeled all four
%methods of the \mEdhoc{} \mSpec{} using \mTamarin.
%
% We formulated several important security properties and identified precise
% adversary models in which we verified these.
% %
%The properties are shown in Table~\ref{tab:props}.
%
% Mutual injective agreement covers the parameters $S_P$:
% responder identity, roles, session-key material (except for $P_I$ when
% initiator uses the \mStat{} authentication
% method), context identifiers \mCi{} and \mCr, and cipher suites \mSuites.
% %
% The responder in addition is ensured agreement on the initiators identity and
% $P_I$, i.e., on the set $S_F$.
% %
% Implicit agreement covers all previously mentioned parameters for both 
% peers.
% %
% All measurements were taken on a server with 2 Intel(R) Xeon(R) Gold
% 6242 CPU @ 2.80GHz with a total of 32 cores / 64 threads and reserved
% 256 GiB of memory for the scheduled job (the machine has a total of
% 384 GiB installed).
% %
% Time measurements are rounded to the nearest second and we report both
% the real time (time to complete the task) and user time (total CPU time
% spent for the task).
% %
Mutual entity authentication, UKS- and KCI-resistance can be inferred
from the verified properties.
%

@@

Further, we identified a situation where initiators may establish an 
\mOscore{}
security context with a different party than the application using \mEdhoc{}
intended, and proposed a simple mitigation.
%
We discussed how the IETF may extract and better define security properties 
to
enable easier verification.
%
We also discussed some aspects of the protocol with respect to error handling and denial of service attacks.

@@

\subsection{A Note About the Version of \mEdhoc}
\label{sec:newdrafts}
In this work, we have analyzed the \mEdhoc{} specification as of July
2020~\cite{our-analysis-selander-lake-edhoc-00}.
%
While our formal analysis applies to v5 of the \mSpec{} (February 2021),
there are newer versions.
%
The latest version at the time of writing is
v15 (July 2022)~\cite{draft-ietf-lake-edhoc-15}, which, among other things,
differ from v5 in terms of some inputs to the key derivations.
%
Modelling the new key derivation function would have needed sweeping changes across our entire formal model, and due to paucity of time, we could not modify the Tamarin model to be up to date with the current version.
%
However, we do refer to various aspects
(error handling, denial of service etc) of this latest version,
in Sections~\ref{sec:errorHandling} and \ref{sec:discussion}.
%
Modelling the latest version of the protocol is also left as future work.
%

>>>>

%-------------------------------------------------------------------------- sub
\subsection{Contributions}
\label{sec:contributions}
In this paper, we extend the work presented in~\cite{Norr21}.
%
We formally analyze the \mEdhoc{} specification as of July
2020~\cite{our-analysis-selander-lake-edhoc-00}, but our formal analysis applies as far as the version of the \mSpec{} from February 2021.
%
We extend the adversary model and \mTamarin{} system models to capture weak
post-compromise security (PCS) and model Trusted Execution Environments
(TEE).
%
We also alter the formal modeling used for encryption under \mXor{}.
%
We formally verify the following properties:
\begin{itemize}
\item Injective agreement
\item Implicit agreement
\item Perfect Forward Secrecy (PFS) for the session-key material
\item Weak post-compromise security for the session-key material
\item Secrecy and integrity of \mADthree
\end{itemize}
%
We follow the definition of weak PCS by~\cite{DBLP:conf/csfw/Cohn-GordonCG16},
which subsumes PFS.
%
Since \mEdhoc{} is still a protocol under development, the latest version of \mEdhoc{} is the one from July 2022~\cite{draft-ietf-lake-edhoc-15}, which contains details which are not covered by our formal model.
%
We do, however, refer to various aspects
(error handling, denial of service etc) of the July 2022 version of the \mSpec{}, namely 
in Sections~\ref{sec:errorHandling} and \ref{sec:discussion}.
%
We also discuss various issues arising due to the use of trusted
execution environments (TEEs), denial of service (DoS) attacks, error handling,
the negotiation of parameters (which the formal model abstracts
away) for establishing the protocol, and other potential attacks and concerns.
%
We have communicated these issues to the developers of the protocol.
%

@@

%-------------------------------------------------------------------------- sub
\subsection{Session Key Material}
\label{sec:sessionKeyMaterial}
As we have seen, \mEdhoc{} establishes some
session-key material, which can be fed into the \mEdhoc{}-Exporter to derive
session keys.
%
This key material is directly influenced by \mGxy{} and a party's secret static
long-term key, if they use \mStat{}.
%
As mentioned earlier, mutual explicit injective agreement cannot be obtained for
$P_{I}$.
%
This might prove to be a problem, and we proposed three potential alternatives
for addressing this to the \mEdhoc{} authors.
%
\begin{itemize}
    \item Include \mIdcredi{} or a hash thereof in the first and
    second messages:
%
    This increases message size and also leaks the initiator's identity.
%
    \item Derive session-key material without using $P_{I}$:
%
    This deviates from the structure of protocols like \mOptls{} from which the
    \mStat{}-based methods draw heavy inspiration.
%
    In those protocols, including $P_{I}$ while computing session-key
    material is crucial to obtaining resistance against the compromise of the
    initiator's ephemeral key.
%
    \item Include a fourth message from the
    responder to the initiator which contains a MAC based on information
    obtained using $P_{I}$:
\end{itemize}
%
While the last option increases the number of messages in the protocol, the
\mEdhoc{} authors decided to go with this option for situations where it is
necessary to provide explicit injective agreement on $P_{I}$.
%

%-------------------------------------------------------------------------- sec
\section{\uppercase{Conclusions and Future Work}}
\label{sec:conclusion}
As part of this paper, we modeled all four authentication methods of the
\mEdhoc{} protocol using the \mTamarin{} verification tool.
%
We used a server with 2 Intel(R) Xeon(R) Gold 6242 CPU @ 2.80GHz, which had a
total of 32 cores / 64 threads, and of the total 384GiB, the server allotted
256GiB of memory for the \mTamarin{} job.
%
We report both the time taken to complete the task (``real time'') and the CPU
time spent on the task (``user time'').
%
Time measurements are rounded to the nearest second.
%
We formulated and verified the following properties in a precise adversary model
with controlled access to TEEs and a precise modeling of \mXor.
%
\begin{itemize}
\item Injective agreement for both $I$ and $R$
\item Implicit agreement for both $I$ and $R$
\item Perfect Forward Secrecy (PFS) for the session-key material
\item Weak post-compromise security for the session-key material
\item Secrecy and integrity of \mADthree{}
\end{itemize}
%
We consider two separate sets of parameters, a partial set ($S_P$) and the full
set ($S_F$).
%
The set $S_P$ consists of the following pieces of information: the identity of
the responder, roles, context identifiers \mCi{} and \mCr, cipher suites
\mSuites, and session-key material excluding $P_I$.
%
For injective agreement, when the initiator uses a \mStat{} method, they are
ensured agreement only on this above set of parameters.
%
The set $S_F$ is the set $S_P$ along with the initiator's identity and $P_I$.
%
The responder is guaranteed injective agreement on the parameters $S_F$,
irrespective of what method they run, while the initiator is guaranteed
injective agreement on $S_F$ as long as they run a \mSig{} method.
%
Implicit agreement can be reached for both parties on the set $S_F$.
%
In addition, mutual entity authentication, UKS- and KCI-resistance can be
inferred from the verified properties.
%
We present the results in Table~\ref{tab:props}.
%

Mutual entity authentication, UKS- and KCI-resistance can be inferred
from the verified properties.
%

@@

Further, we identified a situation where initiators may establish an \mOscore{}
security context with a different party than the application using \mEdhoc{}
intended, and proposed a simple mitigation.
%
We discussed how the IETF may extract and better define security properties
to enable easier verification.
%
We also discussed some aspects of the protocol with respect to error handling
and denial of service attacks.
%

@@

\subsection{A Note About the Version of \mEdhoc}
\label{sec:newdrafts}
In this work, we have analyzed the \mEdhoc{} specification as of July
2020~\cite{our-analysis-selander-lake-edhoc-00}.
%
While our formal analysis applies to v5 of the \mSpec{} (February 2021),
there are newer versions.
%
The latest version at the time of writing is
v15 (July 2022)~\cite{draft-ietf-lake-edhoc-15}, which, among other things,
differ from v5 in terms of some inputs to the key derivations.
%
Modeling the new key derivation function would have needed sweeping changes
across our entire formal model, and due to paucity of time, we could not modify
the \mTamarin{} model to be up to date with the current version.
%
However, we do refer to various aspects
(error handling, denial of service etc) of this latest version,
in Sections~\ref{sec:errorHandling} and \ref{sec:discussion}.
%
Modeling the latest version of the protocol is also left as future work.
%
